---
title: "Lab 5"
output:
  github_document:
  html_document:
    html_preview: false
always_allow_html: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part1 Setup the Git project and the GitHub repository

```{r packages}
library(data.table)
library(tidyverse)
library(dplyr)
library(dtplyr)
```
```{r data-read, cache=TRUE}
#cache=true store the content and make running faster
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")

```
#Setup in R
##Load the data.table (and the dtplyr and dplyr packages if you plan to work with those).

##Load the met data from https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz, and also the station data. For the later, you can use the code we used during lecture to pre-process the stations data:
```{r}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```
##Merge the data as we did during the lecture.

```{r}
met <- merge(
  x = met,
  y = stations,
  all.x = TRUE, all.y = FALSE,
  by.x = "USAFID", by.y = "USAF"
) 
```


###Question 1: Representative station for the US
#What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?

#Knit the document, commit your changes, and Save it on GitHub. Don’t forget to add README.md to the tree, the first time you render it.

```{r bcollapsing-by-station}
station_averages <- met[,.(
  temp      = mean(temp, na.rm = TRUE),
  wind.sp   = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE)
), by = .(USAFID)]

```
Now, we need to identify the median per variable
```{r quantiles}
medians <- station_averages[,.(
  temp_50      = quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50   = quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = .5, na.rm = TRUE)
)]

medians

```

Now, we can find the stations that are the closest the these. 

```{r median_temp}
station_averages[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station
```
The median temperature station is `r median_temp_station$USAFID`. 

###Question 2: Representative station per state
#Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.

#Knit the doc and save it on GitHub.

we first need to recover the stae variable, by merging!
```{r}
station_averages <- merge(
  x = station_averages, y = stations,
  by.x = "USAFID", by.y = "USAF",
  all.x = TRUE, all.y = FALSE
  )

```

now we can compute the median per state
```{r}
station_averages[, temp_50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, wind.sp_50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE]


```

Now, the euclidean distance... $\sqrt{\sum_i(x_i - y_i)^2}$

```{r}
station_averages[, eudist := sqrt(
  (temp - temp_50)^2 + (wind.sp - wind.sp_50)^2
  )]
station_averages

```


###Question 3: In the middle?
#For each state, identify what is the station that is closest to the mid-point of the state. Combining these with the stations you identified in the previous question, use leaflet() to visualize all ~100 points in the same figure, applying different colors for those identified in this question.

#Knit the doc and save it on GitHub.

###Question4 

```{r}
met[, state_temp := mean(temp, na.rm = TRUE), by = STATE]
met[, temp_cat := fifelse(
  state_temp < 20, "low-temp", 
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]

```
let's make sure that we don't have NAs
```{r}
table(met$temp_cat, useNA = "always")

```

now, let's summarize
```{r}
tab <- met[, .(
  N_entries  = .N,
  N_stations = length(unique(USAFID))
), by = temp_cat]

knitr::kable(tab)

```


